```{r global_setup, echo=FALSE, cache=FALSE, warning=FALSE}

    opts_chunk$set(cache=FALSE, cache.path='cache/clickme_tutorial/', fig.path='figure/clickme_tutorial/', autodep=TRUE, fig.cap="", fig.width=8, fig.height=8, dev = c("svg", "cairo_pdf"), echo=FALSE, warning=FALSE, message=FALSE)
    dep_auto()

```

# Clickme Points Example

```{r  results="asis"}

    library(clickme)
    data(microarray)
    clickme(points, x = microarray$significance, y = microarray$logFC,
            color_groups = ifelse(microarray$adj.P.Val < 1e-4, "Significant", "Noise"),
            names = microarray$gene_name,
            xlab = "Significance (-log10)", ylab = "Fold-change (log2)",
            extra = list(Probe = microarray$probe_name),
            dir = "/Users/nacho/Documents/Blog/nachocab.github.io/clickme/clickme_tutorial", file = "tutorial2-points")$iframe(height = 760)

```


Things you can do with the chart above:

* Search by gene name
* Zoom in and out using your mouse wheel or trackpad.
* Click and drag the whitespace between points to move around.
* Click on `Show names` to toggle the name of each point.
* Hover over individual points to see additional information.
* Click on the groups to show/hide them.

I’m still working on the documentation. If you want to learn more, check out the [source code](https://github.com/nachocab/clickme).

## Code

You can reproduce the example by running the following in R

```r
# you don't need to run this line if you already have the devtools package installed.
install.packages("devtools")

# every time you run these two lines you update to the latest version of Clickme. Feel free to do so as often as you like.
library(devtools)
install_github("clickme", "nachocab")

# this line loads Clickme. You can put it in your .Rprofile so it is always ready to use.
library(clickme)
data(microarray)
clickme(points, x = microarray$significance, y = microarray$logFC,
        color_groups = ifelse(microarray$adj.P.Val < 1e-4, "Significant", "Noise"),
        names = microarray$gene_name,
        xlab = "Significance (-log10)", ylab = "Fold-change (log2)",
        extra = list(Probe = microarray$probe_name))

```
